Alignment of Light and Mass in Lensing Galaxies

### 2013-05-13 shapes/paper.tex
update on paper

% Open Tasks
%
% --- Keywords ---
% - Add keaywords
%
% --- Introduction ---
% - Citiations needed in parametric vs. nonparametric comparison part
% - Citiations needed on effect of environment
% - Section on how this work might be able to rule out certain alternative gravity models
% - Describe theoretical expectations (ellipticities of dm halo and luminous part, misalignments, effect of environment)
% - ...
%
%
% --- Data ---
% - More detailed description how the baryionic mass distribution was measured
% - ...
%
%
% --- Method ---
% - Maybe more details on GLASS and how it works (cite Jonathan's upcoming paper)
% - Plot showing how well the shape informations is recovered when applied to fake data
% - ...
%
%
% --- Results ---
% - Interpretation of 
% - Interpretation of ellipticities plot (why it is relevant that dm halos are significantly rounder)
% - Interpretation of misalignment versus ellipticity plot ('forbidden region', alignment with semi-minor axis, very elliptical galaxies well aligned)

@@

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Somewhere in France, France
}

@@

\subsection{Context/Background/Motivation}
\begin{itemize}
\item Why interesting (lensing can constrain mass distribution (enclosed mass constraint independent from model with only few percent uncertainty e.g. Kochanek 1991?, Wambsganss \& Paczyński 1994?), insights into structure formation)
\item What is to be expected from theory (dm, alternative gravity)
\end{itemize}

\subsection{Anything done before?}
\begin{itemize}
\item Surely cite: \cite{1997ApJ...482..604K,1998ApJ...509..561K,2006ApJ...649..599K,2009ApJ...690..670T,2012ApJ...761..170G}
\end{itemize}

\subsection{New here}
\begin{itemize}
\item Statistics
\item Free-form modelling
\item group versus stellar misalignment with dm halo
\end{itemize}

@@

\begin{table*}
 \begin{center}
  \begin{tabular}{l c c c c c c c}
   \multirow{2}{*}{Lens} & \multirow{2}{*}{Environment$^{a}$} & \multicolumn{2}{c}{Group/Cluster centroid$^{b}$} & & \multicolumn{2}{c}{Other galaxy} & \multirow{2}{*}{References$^{c}$}\\ \cline{3-4} \cline{6-7}
    & & $\Delta\alpha$ [''] & $\Delta\delta$ [''] & & $\Delta\alpha$ [''] & $\Delta\delta$ ['']\\ \hline \hline
   Q0047 & G(9) & $3.6\pm15.6$ & $43.2\pm28.8$ &   & ... & ... & 1\\
   MG0414 & 2 & ... & ... &   & $-0.4$ & $1.5$ & 2\\
   B0712 & 2 & ... & ... &   & $47$ & $-91$ & 3\\
   RXJ0911 & C & $-12.0$ & $-39.8$ &    & ... & ... & 4\\
   Q0957 & C & ... & ... &   & ... & ... & 5\\
   B1030 & 2 & ... & ... &   & $-9.8$ & $8.5$ & 6\\
   PG1115 & G(13) & $-10.8\pm21.0$ & $-3.6\pm15.6$ &   & ... & ... & 1\\
   B1422 & G(17) & $61.2\pm25.2$ & $-10.8\pm23.4$ &   & ... & ... & 1\\
%   SBS1520 & G(5) & ... & ... &   & ... & ... & 7\\
%   B1600 & G(17) & $-8.4^{d}$ & $-1.5^{d}$ &   & ... & ... & 8\\
   B1608 & G(8) & $-0.7$ & $8.0$ &   & ... & ... & 9\\
   MG2016 & C(69) & ... & ... &   & ... & ... & 10\\
   HE2149 & G & $-36.6$ & $14.2$ &   & ... & ... & 11\\
  \end{tabular}
  \caption{List of lenses lying in a group or cluster environment, or that have a companion. \newline $^{a}$ G stands for group, C for cluster, and 2 means that the lens has a companion galaxy. If known, the number of member galaxies including the lens is given. \newline $^{b}$ If not stated differently, the centroid positions are luminosity-weighted estimates. \newline $^{c}$ References for the centroid estimates we used. \newline $^{d}$ Only number-weighted estimates are available. \newline 1: \cite{2011ApJ...726...84W}; 2: \cite{1993AJ....105....1S}; 3: \cite{2002AJ....123..627F}; 4: \cite{2001ApJ...555....1M}; 5: \cite{1998ApJ...504..661C}; 6: \cite{2000ApJ...536..584L}; 7: \cite{2008ApJ...673..778A}; 8: \cite{2007AJ....134..668A}; 9: \cite{2003MNRAS.344..337T}; 10: \cite{2006ApJ...646...85W}; 11: \cite{2006ApJ...646...85W}}
  \label{tab:environment}
 \end{center}
\end{table*}

@@

Our sample originally consisted of 21 galaxies. The sample is identical to the one as in \cite{leier11phd} and \cite{2011ApJ...740...97L}. As the lenses \textit{HS0818+1227}, \textit{SBS1520+530}, \textit{B1600+434}, \textit{B1608+656} proved too difficult to reconstruct, we dropped them from this analysis. The final sample consists thus of 17 galaxies. Most of these lenses are early-type galaxies except \textit{B1030+074}, \textit{B1152+200}, and \textit{Q2237+030}. All lenses can also be found in the gravitational lens database \textit{CASTLES}. Characteristics and references of each lens in the sample are listed below. Table \ref{tab:overview} contains all the redshifts and the type of the lenses, Table \ref{tab:environment} lenses with environment information, and Table \ref{tab:timedelays} lenses with measured time delays.

@@

\textit{RXJ0911+0551} (hereafter \textit{RXJ0911}) is an almost circular early-type galaxy \citep{2012A&A...538A..99S}. It has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. This cluster seems to be rather complex and especially it is not spherical. Possibly it is not dynamically relaxed, although X-ray emission can be detected. Analysis of this emission yields a temperature of 2.3 keV. There is also a satellite galaxy to the north-west direction \citep{2000ApJ...544L..35K}.

@@

\textit{Q0957+561} (hereafter \textit{Q0957}) is a cD galaxy lying close to the center of a cluster with a high sprial galaxy-fraction (e.g. \cite{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}). Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the center of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. These results are consistent with the center of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay, e.g. \citep{2012A&A...540A.132S}. They however also find a three-day lag between the g- and r-bands, and the estimates do not agree at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

@@

\textit{HE1104-1805} (hereafter \textit{HE1104}) is probably an early-type galaxy \citep{2000ApJ...536..584L,2000A&A...360..853C}. It has measured time delays \citep{2008ApJ...676...80M}. \cite{2000ApJ...536..584L} also seem to find a group environment. \cite{2004A&A...428..741F} however conclude that the photometric redshift measurements of these galaxies indicate that they are at a higher redshift and rather associated with the source. Nonetheless, their lens models, parametric and free-form, all require strong external shear.

@@

\textit{B1422+231} (hereafter \textit{B1422}) is an early-type galaxy \citep{1996ApJ...462L..53I}. Although there are measured time delays by \cite{2001MNRAS.326.1403P}, they are probably not to be trusted too much as they deviate significantly from theoretical expectations in \cite{2003AJ....126...29R}. An accurate measurement of the time delays would require time delay measurements on the scale of hours. \cite{2006ApJ...641..169M} find a group environment with 16 spectroscopically confirmed member galaxies. Using newer data, \cite{2011ApJ...726...84W} find an additional member. \cite{2004ApJ...610..686G} also detect X-ray emission from the corresponding group at a temperature of 1.0 keV.

@@

\textit{MG2016+112} (hereafter \textit{MG2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster which consists of 69 probable galaxies \citep{2003MNRAS.344..337T}. The clusters shows a high density of galaxies close to the lens in a south-east direction.

@@

\textit{HE2149-2745} (hereafter \textit{HE2149}) is an early-type galaxy \citep{2007A&A...465...51E}. It has a measured time delay \citep{2002A&A...383...71B}. There is some uncertainty in estimating the lens' redshift. Initial estimates by \cite{1996A&A...315L.405W,2000ApJ...543..131K} found a probable redshift range of $\sim0.2-0.5$. By cross-correlating the lens spectrum with a template spectrum of an elliptical galaxy, \cite{2002A&A...383...71B} find that the redshift probably lies in the range $0.49\leq z \leq 0.60$, with the most likely redshift $z=0.495\pm0.01$. \cite{2007A&A...465...51E} however find that the redshift is more likely to be $z=0.603\pm0.001$. This would probably make the lens part of a poor group found by \cite{2006ApJ...641..169M} and \cite{2006ApJ...646...85W}. They also find two other groups at lower redshifts.

\textit{Q2237+030} (hereafter \textit{Q2237} or \textit{2237}) is a barred spiral \citep{1988AJ.....95.1331Y}. With only a redshift of $z=0.04$, it is the closest lens of the sample. Due to the low redshift, the probed physical scales are quite small. No further inquiries into the environment were made.

>>>>

% Open Tasks
%
% --- General ---
% - Add keywords
% - Check universities' names
% - Correct typos and reformulate some paragraphs
% - ...
%
%
% --- Introduction ---
% - Citiations needed in parametric vs. nonparametric comparison part
% - Citiations needed on effect of environment
% - Section on how this work might be able to rule out certain alternative gravity models
% - Describe theoretical expectations (ellipticities of dm halo and luminous part, misalignments, effect of environment)
% - ...
%
%
% --- Data ---
% - More detailed description how the baryionic mass distribution was measured
% - ...
%
%
% --- Method ---
% - Maybe more details on GLASS and how it works (cite Jonathan's upcoming paper)
% - Plot showing how well the shape informations is recovered when applied to fake data
% - ...
%
%
% --- Results ---
% - Interpretation of 
% - Interpretation of ellipticities plot (why it is relevant that dm halos are significantly rounder)
% - Interpretation of misalignment versus ellipticity plot ('forbidden region', alignment with semi-minor axis, very elliptical galaxies well aligned)

@@

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Department of Biology and Health, Versailles Saint-Quentin-en-Yvelines University, France
}

@@

\textbf{Context/Background/Motivation}
\begin{itemize}
\item Why interesting (lensing can constrain mass distribution (enclosed mass constraint independent from model with only few percent uncertainty e.g. Kochanek 1991?, Wambsganss \& Paczyński 1994?), insights into structure formation)
\item What is to be expected from theory (dm, alternative gravity)
\end{itemize}

\textbf{Anything done before?}
\begin{itemize}
\item Surely cite: \cite{1997ApJ...482..604K,1998ApJ...509..561K,2006ApJ...649..599K,2009ApJ...690..670T,2012ApJ...761..170G}
\end{itemize}

\textbf{New here}
\begin{itemize}
\item Statistics
\item Free-form modelling
\item group versus stellar misalignment with dm halo
\end{itemize}

@@

\begin{table*}
 \begin{center}
  \begin{tabular}{l c c c c c c c}
   \multirow{2}{*}{Lens} & \multirow{2}{*}{Environment$^{a}$} & \multicolumn{2}{c}{Group/Cluster centroid$^{b}$} & & \multicolumn{2}{c}{Other galaxy} & \multirow{2}{*}{References$^{c}$}\\ \cline{3-4} \cline{6-7}
    & & $\Delta\alpha$ [''] & $\Delta\delta$ [''] & & $\Delta\alpha$ [''] & $\Delta\delta$ ['']\\ \hline \hline
   Q0047 & G(9) & $3.6\pm15.6$ & $43.2\pm28.8$ &   & ... & ... & 1\\
   MG0414 & 2 & ... & ... &   & $-0.4$ & $1.5$ & 2\\
   B0712 & 2 & ... & ... &   & $47$ & $-91$ & 3\\
   RXJ0911 & C & $-12.0$ & $-39.8$ &    & ... & ... & 4\\
   Q0957 & C & ... & ... &   & ... & ... & 5\\
   B1030 & 2 & ... & ... &   & $-9.8$ & $8.5$ & 6\\
   PG1115 & G(13) & $-10.8\pm21.0$ & $-3.6\pm15.6$ &   & ... & ... & 1\\
   B1422 & G(17) & $61.2\pm25.2$ & $-10.8\pm23.4$ &   & ... & ... & 1\\
%   SBS1520 & G(5) & ... & ... &   & ... & ... & 7\\
%   B1600 & G(17) & $-8.4^{d}$ & $-1.5^{d}$ &   & ... & ... & 8\\
   B1608 & G(8) & $-0.7$ & $8.0$ &   & ... & ... & 9\\
   MG2016 & C(69) & ... & ... &   & ... & ... & 10\\
   HE2149 & G & $-36.6$ & $14.2$ &   & ... & ... & 11\\
  \end{tabular}
  %$^{d}$ Only number-weighted estimates are available. \newline 
  \caption{List of lenses lying in a group or cluster environment, or that have a companion. \newline $^{a}$ G stands for group, C for cluster, and 2 means that the lens has a companion galaxy. If known, the number of member galaxies including the lens is given. \newline $^{b}$ If not stated differently, the centroid positions are luminosity-weighted estimates. \newline $^{c}$ References for the centroid estimates we used. \newline 1: \cite{2011ApJ...726...84W}; 2: \cite{1993AJ....105....1S}; 3: \cite{2002AJ....123..627F}; 4: \cite{2001ApJ...555....1M}; 5: \cite{1998ApJ...504..661C}; 6: \cite{2000ApJ...536..584L}; 7: \cite{2008ApJ...673..778A}; 8: \cite{2007AJ....134..668A}; 9: \cite{2003MNRAS.344..337T}; 10: \cite{2006ApJ...646...85W}; 11: \cite{2006ApJ...646...85W}}
  \label{tab:environment}
 \end{center}
\end{table*}

@@

Our sample originally consisted of 21 galaxies, identical to the one in \cite{leier11phd} and \cite{2011ApJ...740...97L}. As the lenses \textit{HS0818+1227}, \textit{SBS1520+530}, \textit{B1600+434}, \textit{B1608+656} proved too difficult to reconstruct, we dropped them from this analysis. The final sample consists thus of 17 galaxies. Most of these lenses are early-type galaxies except \textit{B1030+074}, \textit{B1152+200}, and \textit{Q2237+030}. All lenses can also be found in the gravitational lens database \textit{CASTLES}. Characteristics and references of each lens in the sample are listed below. Table \ref{tab:overview} contains all the redshifts and the type of the lenses, Table \ref{tab:environment} lenses with environment information, and Table \ref{tab:timedelays} lenses with measured time delays.

@@

\textit{RXJ0911+0551} (hereafter \textit{RXJ0911}) is an almost circular early-type galaxy \citep{2012A&A...538A..99S}. It has measured time delays \citep{2002ApJ...572L..11H} and lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. This cluster seems to be rather complex. Possibly it is not dynamically relaxed, although X-ray emission can be detected. Analysis of this emission yields a temperature of 2.3 keV. There is also a satellite galaxy to the north-west direction \citep{2000ApJ...544L..35K}.

@@

\textit{Q0957+561} (hereafter \textit{Q0957}) is a cD galaxy lying close to the center of a cluster with a high sprial galaxy-fraction (e.g. \cite{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}). Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the center of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. These results are consistent with the center of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay (e.g. \cite{2012A&A...540A.132S}). They however also find a three-day lag between the g- and r-bands, and the estimates do not agree at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

@@

\textit{HE1104-1805} (hereafter \textit{HE1104}) is probably an early-type galaxy \citep{2000ApJ...536..584L,2000A&A...360..853C} and has measured time delays \citep{2008ApJ...676...80M}. \cite{2000ApJ...536..584L} also seem to find a group environment. \cite{2004A&A...428..741F} however conclude that the photometric redshift measurements of these galaxies indicate that they are at a higher redshift and rather associated with the source. Nonetheless, their lens models, parametric and free-form, all require strong external shear.

@@

\textit{B1422+231} (hereafter \textit{B1422}) is an early-type galaxy \citep{1996ApJ...462L..53I}. Although there are measured time delays by \cite{2001MNRAS.326.1403P}, they are probably not to be trusted too much as they deviate significantly from theoretical expectations in \cite{2003AJ....126...29R}. An accurate measurement of the time delays would require time delay measurements on the scale of hours. \cite{2006ApJ...641..169M} find a group environment with 16 spectroscopically confirmed member galaxies. Using newer data \cite{2011ApJ...726...84W} find an additional member. \cite{2004ApJ...610..686G} also detect X-ray emission from the corresponding group at a temperature of 1.0 keV.

@@

\textit{MG2016+112} (hereafter \textit{MG2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster which consists of 69 probable member galaxies \citep{2003MNRAS.344..337T}. The clusters shows a high density of galaxies close to the lens in a south-east direction.

@@

\textit{HE2149-2745} (hereafter \textit{HE2149}) is an early-type galaxy \citep{2007A&A...465...51E} and has a measured time delay \citep{2002A&A...383...71B}. There is some uncertainty in estimating the lens' redshift. Initial estimates by \cite{1996A&A...315L.405W} and \cite{2000ApJ...543..131K} found a probable redshift range of $\sim0.2-0.5$. By cross-correlating the lens spectrum with a template spectrum of an elliptical galaxy, \cite{2002A&A...383...71B} find that the redshift probably lies in the range $0.49\leq z \leq 0.60$, with the most likely redshift $z=0.495\pm0.01$. \cite{2007A&A...465...51E} however find that the redshift is more likely to be $z=0.603\pm0.001$. This would probably make the lens part of a poor group found by \cite{2006ApJ...641..169M} and \cite{2006ApJ...646...85W}. They also find two other groups at lower redshifts.

\textit{Q2237+030} (hereafter \textit{Q2237} or \textit{2237}) is a barred spiral \citep{1988AJ.....95.1331Y}. With only a redshift of $z=0.04$ it is the closest lens of the sample. Due to the low redshift, the probed physical scales are quite small. No further inquiries into the environment were made.
